Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest

### 2020-04-02 Manuscript.Rmd
20200402

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Sp, -LTD, -SHA)
suppressWarnings(row.names(PFT) <- PFT$SpCode) # deprecated
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
Q <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(data.frame(PFT.pca$scores) %>% 
              mutate(SpCode = row.names(.)),
            by = "SpCode") %>% 
  rename(WES = Comp.1, LES = Comp.2) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  select(SLA, LDMC, WD, LA, LT, LES, WES, strata, SpCode) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata))
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
L <- genTrees() %>% 
  mutate(com = (XY %over% com)[,1]) %>% 
  filter(!is.na(com)) %>% 
  group_by(com, SpCode) %>% 
  summarise(n = n()) %>% 
  dcast(com ~ SpCode, value.var = "n") %>% 
  replace(is.na(.), 0)
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWMses
GWM <- L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(Q, by = "SpCode") %>% 
  filter(!is.na(strata)) %>% 
  select(-SpCode) %>% 
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>% 
  group_by(strata, community) %>% 
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[which(x[,"strata"] == "C"),'SpCode'] <- sample(x[which(x[,"strata"] == "C"),'SpCode'])
#   x[which(x[,"strata"] == "U"),'SpCode'] <- sample(x[which(x[,"strata"] == "U"),'SpCode'])
#   return(x)
#   }), c(dim(Q), n), list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(detectCores())
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     melt(id.var = "com",
#          variable.name = "SpCode",
#          value.name = "abundance") %>%
#     dcast(SpCode ~ com) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path("draft_save", 'nGWM.Rdata'))
load(file.path("draft_save", 'nGWM.Rdata'))
GWM_ses <- GWM
GWM_ses[3:9] <- (GWM[3:9] - apply(nGWM, 1:2, mean, na.rm = T))/apply(nGWM, 1:2, sd, na.rm = T)
```

@@

```{r LM}
# GWM_ses %>% 
#   filter(!is.na(strata)) %>% 
#   melt(id.vars = c('strata', 'community')) %>% 
#   rename(trait = variable) %>% 
#   rename(SES = value) %>% 
#   mutate(community = as.character(community)) %>% 
#   left_join(R, by = "community") %>%
#   group_by(strata, trait) %>% 
#   do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#      direction = 'both', trace = 0)) %>% 
#   # do(reg = lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .),
#   #    direction = 'both', trace = 0) %>%
#   tidy(reg) %>%
#   group_by(strata, trait, term) %>%
#   # transmute(value = paste(format(estimate, scientific = T, digits = 2, trim = T), 
#                           # stars(p.value, ns = '   '))) %>% 
#   transmute(value = format(estimate, scientific = T, digits = 2, trim = T)) %>% 
#   dcast(strata + trait ~ term) %>% 
#   select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>% 
#   mutate_all(funs(replace(., is.na(.), ' '))) %>% 
#   left_join( # Inclusion of R2 but should be split
#     GWM_ses %>% 
#       filter(!is.na(strata)) %>% 
#       melt(id.vars = c('strata', 'community')) %>% 
#       rename(trait = variable) %>% 
#       rename(SES = value) %>% 
#       mutate(communtiy = as.character(community)) %>% 
#       left_join(R, by = "community") %>%
#       group_by(strata, trait) %>% 
#       do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
#                     direction = 'both', trace = 0)) %>% 
#       glance(reg) %>% 
#       select(r.squared) %>% 
#       mutate(r.squared = round(r.squared, 2)) %>% 
#       rename("$R^2$" = r.squared), by = c("strata", "trait")
#   )%>% # End of R2
#   filter(trait != 'LES') %>% 
#   filter(trait != 'WES') %>%
#   # bold() %>% 
# Coefficients are unstable depending on random GWM so I entered the one from the last online version
data.frame(
  Guild = c(rep("C", 5), rep("U", 5)), Trait = rep(c("SLA", "WD", "LDMC", "LA", "LT"), 2),
  Slope = c(-9.5*10^-3, "", "", "", 1.3*10^-2, "", 1.3*10^-2, "", "", ""),
  Curvature = c(-2, 3.8, -3.2, -6, 3, -1.3, 6, 2.9, "", 6.6),
  Wetness = c(-0.16, "",  -0.28, "", "", -0.22, "", "", 0.32, 0.19),
  SW = c("", 0.23, "", -0.19, "", "", "", "", "", ""),
  Canopy = c("", "", "", 1.8*10^-2, "", 2.6*10^-2, -1.5*10^-2, -9*10^-3, "", -1.6*10^-2),
  R2 =c(0.06, 0.11, 0.04, 0.18, 0.05, 0.5, 0.32, 0.06, 0.04, 0.19)
) %>% 
  rename("Adjusted $R^2$" = R2) %>% 
  kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$ . See Table \\@ref(tab:traits) for trait abbreviations.', format = 'pandoc')
```

>>>>

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n()) %>% 
  filter(community != "C99")
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
  geom_density(alpha = .5) +
  facet_wrap(~ strata + trait, scales = "free") +
  scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, SES, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
```

@@

```{r LM}
group_by(SES, strata, trait) %>% 
  mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), scale, center = F) %>%
  do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                direction = 'both', trace = 0)) %>%
  tidy(reg) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(estimate = round(estimate, 2)) %>% 
  group_by(strata, trait, term) %>%
  dcast(strata + trait ~ term, value.var = "estimate") %>%
  select(strata, trait, Slope, Curvature, Wetness, SW, Canopy) %>%
    mutate_all(funs(replace(., is.na(.), ' '))) %>%
  left_join(group_by(SES, strata, trait) %>% 
              mutate_at(c("Slope", "Curvature", "Wetness", "SW", "Canopy"), 
                        scale, center = F) %>%
              do(reg = step(lm(SES ~ Slope + Curvature + Wetness + SW + Canopy, .), 
                            direction = 'both', trace = 0)) %>%
              glance(reg) %>%
              dplyr::select(r.squared) %>%
              mutate(r.squared = round(r.squared, 2)), 
            by = c("strata", "trait")) %>%
            rename("$R^2$" = r.squared, Guild = strata, Trait = trait) %>%  
    kable(caption = 'Variation of standardized effect size (SES) values, for canopy (C) and understory (U) guilds, depending on slope, curvature, wetness, southwesterness (SW), and canopy height (Canopy). Linear model coefficients are reported only for predictors retained after bidirectional stepwise selection, with corresponding adjusted $R^2$ . See Table \\@ref(tab:traits) for trait abbreviations.', format = 'pandoc')
```
